---
title: "scATAC_seurat_1"
author: "Irzam Sarfraz"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
.libPaths("/projectnb/paxlab/isarfraz/RProjects/libs")

library(ggplot2)
library(Seurat)
library(Signac)
library(EnsDb.Hsapiens.v86)
library(cowplot)
library(SingleCellExperiment)
```
# Script 3

1. Read in peaks x cells matrix and convert to seurat
NOTE: Here we only need healthy all cells

```{r}
atac_hl <- readRDS("/projectnb/paxlab/isarfraz/Data/Seurat_Final_Data/scATAC-Healthy-Hematopoiesis-191120.rds")
print(dim(atac_hl))
# read in previously computed peaks x cells
# rows are peaks and columns are cells

peaks <- assay(atac_hl, "counts")
chrom_assay <- CreateChromatinAssay(counts = peaks, sep = c("_", "_"))
atac_hl_srt <- CreateSeuratObject(
  counts = chrom_assay,
  assay = "peaks"
)
```

2. Normalization and dimensionality reduction using LSI (TF-IDF + SVD)
```{r}
atac_hl_srt <- RunTFIDF(atac_hl_srt)
atac_hl_srt <- RunSVD(atac_hl_srt, n = 51, features = rownames(atac_hl_srt))
# using all features at this point
DepthCor(atac_hl_srt)
# first componenet is correlated, don't use it downstream
```

3. Clustering using SVD components (SNN)
```{r}
atac_hl_srt <- FindNeighbors(object = atac_hl_srt, dims = 2:51, reduction = "lsi")
atac_hl_srt <- FindClusters(object = atac_hl_srt, resolution = 1.5)
atac_hl_srt <- RunUMAP(object = atac_hl_srt, reduction = 'lsi', dims = 2:51) # default params

DimPlot(object = atac_hl_srt, reduction = "umap", label = TRUE)
```

4. Find 50k variable peaks (this seurat method is using percetange of counts as cutoff instead of variance which greenleaf paper uses) and subset re-process
```{r}
top50k_peaks <- rownames(Signac::FindTopFeatures(atac_hl_srt[['peaks']][]))[1:50000]
head(top50k_peaks)

top50k_atac_hl_srt <- atac_hl_srt[top50k_peaks, ]

top50k_atac_hl_srt <- RunTFIDF(top50k_atac_hl_srt)
top50k_atac_hl_srt <- RunSVD(top50k_atac_hl_srt, n = 51, features = rownames(top50k_atac_hl_srt))

top50k_atac_hl_srt <- FindNeighbors(object = top50k_atac_hl_srt, dims = 2:51, reduction = "lsi")
top50k_atac_hl_srt <- FindClusters(object = top50k_atac_hl_srt, resolution = 1.5)
top50k_atac_hl_srt <- RunUMAP(object = top50k_atac_hl_srt, reduction = 'lsi', dims = 2:51, min.dist = 0.001) #vary mindist
# subset to top50k peaks
# renormalize
# re-cluster

# top50k_atac_hl_srt <- RunUMAP(object = top50k_atac_hl_srt, reduction = 'lsi', dims = 2:51, n.neighbors = 55, min.dist = 0.45, metric = "euclidean") # these are from greenleaf but makes bad umap because of different process and top features


DimPlot(object = top50k_atac_hl_srt, reduction = "umap", label = TRUE)
```

# Script 4

1. Compute gene scores (gene activity matrix)
needs fragments tsv files
```{r}
fpath <- "/projectnb/paxlab/isarfraz/Data/Seurat_Final_Data/GSM4138888_scATAC_BMMC_D5T1.fragments.tsv.gz"
# tell here what to do if you dont have tabix file
# install samtools then
# https://github.com/stuart-lab/signac/issues/242#issuecomment-693666737
fragments <- CreateFragmentObject(
  path = fpath,
  cells = colnames(atac_hl_srt),
  validate.fragments = TRUE
)
#> Computing hash
Fragments(atac_small) <- fragments

```

